\singlespacing \vspace{-2.5cm} \LARGE \textbf{Trade-offs among plant floral and reproductive traits determine interactions with floral visitors}

### 2021-08-16 Supplementary_Material/Supplementary_Material_LK.Rmd
proof-reading LK

Supplementary Table S2: Loadings of the first three axes of trait variation of the phylogenetic informed principal component analysis with the full set of species

Supplementary Table S3: Loadings of the first three axes of trait variation of the phylogenetic informed principal component analysis with the subset of species with data of nectar and pollen quantity

Supplementary Table S4: Statistical association between the different categorical variables and the fisrt three main axes of trait variation with the full set of species

@@

Supplementary Fig. S2: Phylogenetic informed principal component analysis for the species that did not have missing values

Supplementary Fig. S3: Fitted posterior estimates of the number of visits made by bees including and exluding honey bees on the main axes of trait variation

@@

Supplementary Fig. S6: Phylogenetic informed principal component analysis for the subset of species with floral rewards.

@@

Supplementary Fig. S8: Visitation rate of the main bee families on the main axes of trait variation

@@

-	Selfing level: We recorded the selfing level of the different species with both quantitative and qualitative data. The qualitative data was divided in four main categories, high selfers, medium selfers, low selfers and none which was for the species that were unable to self-pollinate. In addition, quantitative data was also divided into these four categories with the following criteria: from 0% to lower than 1% ‘none’, from 1% to 25% ‘low’, from 26% to 75% ‘medium’ and from 76% to 100%  ‘high’.

@@

-	Flower and inflorescence size: We searched for flower length and flower and inflorescence width (mm) for all species. When possible, we calculated the not found measurements with the help of online images and  the software ImageJ.

-	Flower number per plant: We compiled information about the number of flowers per plant for all species. However, this field was rare to find and we also used online images of the different species in order to calculate rough numbers of flowers per plant. We referenced all the filled fields in order to be able to follow the images that were used for these counts. It is important to note, that these numbers are not pretended to be the exact number of flowers per species but an approximate indicator of the reproductive investment for the different species that allow the macroecological analysis of this field.

-	Ovule number: We searched for the number of ovules per flower for all the different species. The number of ovules of Asteraceae species are considered as the total number of ovules per capitulum. Many species were filled by genus or family level because the number of ovules in this taxonomic groups is considered to be mostly constant (e.g., Lamiaceae Boraginaceae or Apiaceae).

-	Style length: The length of the style (mm) of the different species was also included. When possible, we calculated style length from images and illustrations on the online floras.

- Nectar provision: We recorded the presence and absence of nectar for all species. In addition, we also searched for microlitres and milligrams of nectar per flower and nectar concentration. In general terms, nectar data was rarely present and we try to minimise the loss of information by filling some species at family level. For instance, Solanaceae species are described as nectarless, and species that belonged to this family were recorded with ‘absence’ of nectar.

- Pollen grains per flower: We recorded pollen grains per flower for all species but this field was rarely described in the literature. When just the pollen:ovule ratio was found we converted it back to pollen grains by multiplying by ovule number.

@@

- Life form, life span and plant height: We divided the different plant species in 4 main categories: herbs, vines, shrubs and trees. Moreover, we also divided the species between short-lived species (annual, biennial and short-lived perennials) and perennial species (long-lived). Finally, we also searched for the average height (m) of the different species and annotated the maximum and minimum height when possible. We conducted the average between the maximum and minimum height to get an approximate average height of the species when the average was not indicated.

@@

 kable(dat, longtable = T, booktabs = T,linesep = "\\addlinespace", escape=FALSE,  caption = "\\textbf{Supplementary Table S2 | Loadings of the first three axes of trait variation of the phylogenetic informed principal component analysis with the full set of species.}") %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 12, full_width=F,position = "center")

@@

kable(dat, longtable = T,"latex", booktabs = T,linesep = "\\addlinespace", escape=FALSE,  caption = "\\textbf{Supplementary Table S3 | Loadings of the first three axes of trait variation of the phylogenetic informed principal component analysis with the subset of species with data of nectar and pollen quantity.}") %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 12, full_width=F,position = "center")

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10,  fig.cap="\\textbf{Supplementary Fig. S2 |  Phylogenetic informed principal component analysis for the species that did not have missing values.}"}
setwd("~/R_Projects/Reproductive Traits")

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10, fig.cap="\\textbf{Supplementary Fig. S6 | Phylogenetic informed principal component analysis for the subset of species with floral rewards.}"}
setwd("~/R_Projects/Reproductive Traits")

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=28,  fig.cap="\\textbf{Supplementary Fig. S8 | Visitation rate of the main bee families on the main axes of trait variation.} Fitted posterior estimates of the visitation rate of Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae across PC1, PC2 and PC3. We omitted in the plotting area points over percentile 95 \\% percent due to extreme outliers."}

>>>>

Supplementary Table S2: Loadings of the first three axes of trait variation of the phylogenetically informed principal component analysis with the full set of species

Supplementary Table S3: Loadings of the first three axes of trait variation of the phylogenetically informed principal component analysis with the subset of species with data of nectar and pollen quantity

Supplementary Table S4: Statistical association between the different categorical variables and the first three main axes of trait variation with the full set of species

@@

Supplementary Fig. S2: Phylogenetically informed principal component analysis for the species that did not have missing values

Supplementary Fig. S3: Fitted posterior estimates of the number of visits made by bees including and excluding honey bees on the main axes of trait variation

@@

Supplementary Fig. S6: Phylogenetically informed principal component analysis for the subset of species with floral rewards.

@@

Supplementary Fig. S8: Visitation rate of the main bee families in relation to the main axes of trait variation

@@

-	Selfing level: We recorded the selfing level of the different species with both quantitative and qualitative data. The qualitative data were divided in four main categories, high selfers, medium selfers, low selfers and none which was for the species that were unable to self-pollinate. In addition, quantitative data were also divided into these four categories with the following criteria: from 0% to lower than 1% ‘none’, from 1% to 25% ‘low’, from 26% to 75% ‘medium’ and from 76% to 100%  ‘high’.

@@

-	Flower and inflorescence size: We searched for flower length and flower and inflorescence width (mm) for all species. When possible, we measured this trait with the help of online images and the software ImageJ.

-	Flower number per plant: We compiled information about the number of flowers per plant for all species. However, this field was rare to find and we also used online images of the different species in order to calculate rough numbers of flowers per plant. We referenced all the filled fields in order to be able to follow the images that were used for these counts. It is important to note, that these numbers are not predicted to be the exact number of flowers per plant in each species but an approximate indicator of the reproductive investment in each species that will allow for its use in macroecological analysis.

-	Ovule number: We searched for the number of ovules per flower for all species. The number of ovules of Asteraceae species are considered as the total number of ovules per capitulum. Many species were filled at either the genus or family level because the number of ovules in this taxonomic groups is considered to be mostly constant (e.g., Lamiaceae, Boraginaceae, or Apiaceae).

-	Style length: The length of the style (mm) of the different species. When possible, we calculated style length from online images and illustrations.

- Nectar provision: We recorded the presence and absence of nectar for all species. In addition, we also searched for microlitres and milligrams of nectar per flower and nectar concentration. In general terms, nectar data was rarely present and we try to minimise the loss of information by filling some species at family level. For instance, Solanaceae species are described as nectarless. Therefore, all Solanaceae species were recorded with having an ‘absence’ of nectar.

- Pollen grains per flower: We recorded pollen grains per flower for all species but this field was rarely described in the literature. When we could only find the pollen:ovule ratio, we converted it back to pollen grains by multiplying by ovule number.

@@

- Life form, life span and plant height: We divided the different plant species in 4 main categories: herbs, vines, shrubs and trees. Moreover, we also divided the species between short-lived species (annual, biennial and short-lived perennials) and perennial species (long-lived). Finally, we also searched for the average height (m) of the different species and annotated the maximum and minimum height when possible. We calculated the average between the maximum and minimum height to get an approximate average height of the species when the average was not indicated.

@@

 kable(dat, longtable = T, booktabs = T,linesep = "\\addlinespace", escape=FALSE,  caption = "\\textbf{Supplementary Table S2 | Loadings of the first three axes of trait variation of the phylogenetically informed principal component analysis with the full set of species.}") %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 12, full_width=F,position = "center")

@@

kable(dat, longtable = T,"latex", booktabs = T,linesep = "\\addlinespace", escape=FALSE,  caption = "\\textbf{Supplementary Table S3 | Loadings of the first three axes of trait variation of the phylogenetically informed principal component analysis with the subset of species with data of nectar and pollen quantity.}") %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 12, full_width=F,position = "center")

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10,  fig.cap="\\textbf{Supplementary Fig. S2 |  Phylogenetically informed principal component analysis for the species that did not have missing values.}"}
setwd("~/R_Projects/Reproductive Traits")

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10, fig.cap="\\textbf{Supplementary Fig. S6 | Phylogenetically informed principal component analysis for the subset of species with floral rewards.}"}
setwd("~/R_Projects/Reproductive Traits")

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=28,  fig.cap="\\textbf{Supplementary Fig. S8 | Visitation rate of the main bee families in relation to the main axes of trait variation.} Fitted posterior estimates of the visitation rate of Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae across PC1, PC2 and PC3. We omitted in the plotting area points over percentile 95 \\% percent due to extreme outliers."}
